---
title: "Chapter 1 Challenge Solution (R)"
format:
  html:
    code-fold: false
---

## Load libraries

```{r}
library(EIAapi)
library(dplyr)
library(lubridate)
library(plotly)
```

## Question 1

Extract from the EIA dashboard the metadata of the San Diego Gas and Electric balancing authority subregion (SUBBA) under California Independent System Operator parent (PARENT). This series is under the `Hourly Demand By Subregion` sub-category.

## Solution

On the API Dashboard select the following route:

-   Electricity
-   Electric Power Operation (Daily and Hourly)
-   Hourly Demand by Subregion

Once the filters loaded, go to facets and you can either use directly the `SUBBA` filter and find the San Diego Gas and Electric balancing authority subregion (out of the 83 series), or select first California Independent System Operator on the `PARENT` filter and then it narrow down the options the four sub-regions under this parent.

If you go with the last option, you should expect to have the following API URL:

```         
https://api.eia.gov/v2/electricity/rto/region-sub-ba-data/data/?frequency=hourly&data[0]=value&facets[subba][]=SDGE&sort[0][column]=period&sort[0][direction]=desc&offset=0&length=5000
```

Which enables you to extract the API route path (as we saw before) - `electricity/rto/region-sub-ba-data/data`

In addition, you can extract from the header the facets argument for the GET request:

``` json
{
    "frequency": "hourly",
    "data": [
        "value"
    ],
    "facets": {
        "parent": [
            "CISO"
        ],
        "subba": [
            "SDGE"
        ]
    },
    "start": null,
    "end": null,
    "sort": [
        {
            "column": "period",
            "direction": "desc"
        }
    ],
    "offset": 0,
    "length": 5000
}
```

The link to the filtered dashboard is available [here](https://www.eia.gov/opendata/browser/electricity/rto/region-sub-ba-data?frequency=hourly&data=value;&facets=subba;&subba=SDGE;&sortColumn=period;&sortDirection=desc;).

## Question 2

Set a GET request to pull observations between Jan 1st and Jan 31st 2024 with R/Python

## Solution

Using the information we pulled from the API dashboard we can set the GET request parameters:

```{r}
api_key <- Sys.getenv("EIA_API_KEY")

api_path <- "electricity/rto/region-sub-ba-data/data"

frequency <- "hourly"

facets <- list(
    parent = "CISO",
    subba = "SDGE"
)

start <- "2024-01-01T01"
end <- "2024-01-31T23"
```

```{r}
df1 <- eia_get(
    api_key = api_key,
    api_path = api_path,
    frequency = frequency,
    facets = facets,
    start = start,
    end = end
)
```

Let's reformat the `period` variable into a time object:

```{r}
df1 <- df1 |>
    mutate(index = ymd_h(period, tz = "UTC")) |>
    select(index, everything()) |>
    arrange(index)


head(df1)
```

and plot the series:

```{r}
plot_ly(data = df1, x = ~index, y = ~value, type = "scatter", mode = "lines")
```

## Question 3

Use the `eia_backfill` function to pull data from Jan 1st 2020 to Feb 1st 2024

## Solution

Let's use the `eia_backfill` function to pull data between Jan 1st 2020 and Feb 1st 2024:

```{r}
start = as.POSIXct("2020-01-01 00:00:00", tz = "UTC")
end = as.POSIXct("2024-02-01 00:00:00", tz = "UTC")
offset = 2000

df2 <- eia_backfill(
    start =start,
    end = end,
    offset = offset,
    api_key = api_key,
    api_path = api_path,
    facets = facets
)

```

```{r}
str(df2)
```

```{r}
plot_ly(data = df2, x = ~time, y = ~value, type = "scatter", mode = "lines")
```
